\chapter*{Acknowledgements}

First of all, I would like to thank my supervisor, Ana Paiva, from the Faculty of Engineering of University of Porto, for her guidance, insight, encouragement and support during the whole masters. Our weekly moments of discussion were indispensable for the success of this work, and specially for keeping me in the right track and focused on what was important (and not what it would be cool to do).

I would also like to thank all my teachers, from the first grade to my last year in college, as without them I wouldn't have had the necessary skills to accomplish this.

My friends’ friendship has been very important to me and so I thank them for being there for me every step of the way. I would like to specially thank the members of NIAEFEUP, with a very special mention for Pedro, for all the talking, all the comfort and teasing, all the help, and the kind words in the darkest moments. Thank you for all the great moments we have spent together, and for all that we have accomplished together.

A special thank to my parents, Maria and Adelino, without whom I would not be the person I am today and would have not achieved all the great things I have in my life. Thank you for all of your love, your concern and specially your loving support. Thank you for being there for me around the clock.

And finally, a special thank to my boyfriend, Diogo. Thank you for your insight in some of my assignments. Thank you for all of your love, your friendship, the comfort you provided me, your understanding, your patience and for every time you have teased me
along the years.

%\vspace{10mm}
%\flushleft{Clara Raquel da Costa e Silva Sacramento}